Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-10-14 isolation_by_coalescence.tex
more rearrangement

\begin{abstract}
\plr{abstract needs editing}
A number of methods commonly used in landscape genetics
use analogies to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical eqiuvalence of effective resistance
to the mean commute time of a random walk.
Here, we describe the differences between commute time
and coalescence time, which more accurately models genetic distances,
and explore the consequences for inference.
We implement a Bayesian method to do inference under both these models,
and find that inference using commute times
when the data derive from coalescence times
can dramatically mislead in the presence of biased gene flow.
We apply the methods to empirical data,
and find evidence for strongly biased gene flow.
\end{abstract}

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop an inference method for movement rates on a landscape
using both as estimated from genetic distances.
As we will see,
although the two methods are conceptually similar,
commute time inference may get the wrong answer if data derive from a coalescent process,
as does real genetic data, especially in the presence of asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{infer_resistance}.
Others methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem --
inference of movement rates, in either a coalescence or commute time framework,
is an ill-conditioned problem,
implying the need for regularization to obtain reliable inference.
This explains the results of \citet{graves2013current},
who found a large area of resistance parameters
that produced equivalently good fits to genetic distance
(i.e., nonidentifiability, analogous to a flat likelihood surface).

@@

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.0 \citep{haller2017},
with individuals living across continuous, two-dimensional space (sometimes with barriers),
from which we recorded genomic data.
We performed three types of simulation:
uniform migration on a continuous square landscape, 
uniform migration on a rectangular landscape with barriers to migration, 
and biased migration on a square landscape.
Each simulation had $10^4$ diploid, hermaphroditic individuals with nonoverlapping generations.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Offspring are dispersed away from the mother using the same kernel,
conditioned to lie within the habitable area.
In each simulation, $\sigma_d$ was chosen 
so that there were an average of 0.25 individuals per $\sigma_d^2$ units of area.
In order to prevent excessive population clumpiness \citep{felsensten1975pain}, 
we also model local competition. 
To do this, we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals;
then, the fitness of each individual is set to $\exp(-x/40)$, 
where $x$ is the sum of their interaction strengths with other individuals.
This reduces fitness for individuals in denser areas.
Individuals are chosen to reproduce proportional to their fitness.
Clumpiness is similar to what is seen in 
Figure \ref{fig:ind_locs_5x3b_1}.

Individuals have 10 pairs of ``chromosomes'' with $2 \times 10^5$ loci each 
and a recombination rate of $10^{-8}$ per generation per locus.
Mutation rate was $10^{-8}$ per locus for the simulation with uniform migration and no barrier,
and $10^{-7}$ for the remaining simulations.
In all cases, the number of mutations was more than adequate to reliably estimate genetic distance.

To compute genetic distances,
geographic space was partitioned into rectangular regions as depicted below.
within each region, 
individuals are sampled uniformly for ``genotyping'' 
from among those within the middle three quarters of each dimension of rectangle,
as shown for one situation in Figure \ref{fig:ind_locs_5x3b_1}.
This protocol was chosen as a compromise between sampling all individuals 
from near the center of each location,
which could result in a large number of close relatives being sampled, 
leading to an underestimate of the mean genetic distance for individuals in that location
and uniform sampling from the whole area of each grid, 
which may result in many sampled individuals being very close to other grid squares.
Mean genetic distances for each pair of regions is then computed as described above
from genome sequence provided by SLiM.
To estimate uncertainty, we computed the standard error of these mean distances
using the minumum number of individuals in each location.

@@

@@

The data we have used thus far are produced ``under the model'',
and should provide an accurate depiction of our method applied to discrete, randomly mating populations
whose connections by migration are known.
However, this is nearly always a rough approximation to reality,
in which organisms are distributed across continuous geography,
and ``populations'' are constructed by necessity, often driven by sampling locations.

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
and observed, empirical mean estimated using a sample of genotyeps
will differ slightly depending on both the individuals sampled (``sampling noise'')
and the stochasticity of the population history itself (``process noise'').
For more discussion of this latter contribution in other contexts, see
\citet{wakeley2012genealogies}, \citet{waples2009modeling}, or \citet{ralph2015empirical}.
In populations living across continuous geography,
both sources of noise can be substantially larger due to randomness in spatial locations.
A hint of this is visible in Figure \ref{fig:ind_locs_5x3b_1},
where we see that the distribution of individuals across the landscape
is fairly patchy.
This might make inference using a discretization of continuous space considerably more difficult.

@@

\begin{figure}
\centering
\includegraphics[scale=.8]{figs/h_val_comp}
\caption{
    Comparison of the values of the genetic distance matrix $C$
    for a $4 \times 4$ discretization of a square landscape with uniform migration
    calculated from various situations.
    All values are plotted against the corresponding values from the $C$ matrix
    calculated using the first set of individuals from the first simulation,
    along with the $y=x$ line.
    Black circles are from a second (non-overlapping) set of individuals from the first simulation,
    red triangles are from the second simulation, 
    and blue plusses are for the third simulation.
    Mean absolute value relative differences between the new values of $C$ for each situation
    and the original values from the first set of individuals from the first simulation are
    1.8\% for the second set of indivuals from the first simulation,
    13.2\% for the second simulation, 
    and 7.3\% for the third simulation.
    } \label{fig:h_val_comp}
\end{figure}

\paragraph{Identifying a barrier}
We now revisit the earlier situation where a landscape has locations that are barriers to gene flow,
as in Section \ref{sec:5x3b}.
The landscape is designed to be a continuous version of this situation.
Since the distribution of offspring locations is truncated at three standard deviations
away from the mother, a sufficiently thick uninhabitable area can completely stop migration
directly across it.
As in \ref{sec:5x3b}, though, the landscape is still relatively well connected.
The layout of the landscape can be seen in Figure \ref{fig:ind_locs_5x3b_1},
with the red bars being the uninhabitable regions that serve as barriers to migration.
Unlike the case of a square landscape with no barriers, 
there is not a natual way to discretize it more coarsely 
with grid squares of equal size and shape,
so a $5 \times 3$ grid will be the only discretization used in this section.
We calculate mean genetic distances using 50 individuals 
from each grid location,
here giving us a total of 750 individuals.

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/ind_locs_5x3b_1}
    \caption{Locations of all and selected individuals on the $5 \times 3$ landscape with barriers
    in the first replicate.
    Locations of individuals used to compute the mean genetic distance matrix are shown in bold points
    whereas the locations of other individuals are shown in smaller points.
    The red bars show the locations of uninhabitable regions of the landscape.
    They are sufficiently thick so that no migration can occur directly across them.}
    \label{fig:ind_locs_5x3b_1}
\end{figure}

\begin{figure}
\centering
     \includegraphics[height=0.3\textheight]{figs/grid_5x3b_1}
     \includegraphics[height=0.3\textheight]{figs/post_dists_5x3b_1_new}
    \caption{
    \textbf{(top)}
    Posterior medians of movement rates for a replicate of the continuous landscape
    with barriers depicted in Figure \ref{fig:ind_locs_5x3b_1}.
    \textbf{(bottom)}
    Boxplots of the same posterior distributions,
    colored red for gene flow rates of edges that cross the barrier.
    } \label{fig:post_dists_5x3b_1}
\end{figure}

The gene flow rates across the barriers are reliably estimated to be very small
%(\emph{maybe reference all 6 supplemental figures here?}),
but there is considerable variation in the other values.
Note, however, that this analysis requires that we guess a reasonable place
to put the divisions between the locations.
For instance, we may wonder whether a river or other geographical feature
restricts gene flow.
We would then discretize the landscape such that individuals on opposite sides of the 
potential obstruction are in separate grid spaces.

\paragraph{Continuous Landscape with Biased Migration}
We also simulated three situations on a continuous landscape that had an effective bias in migration.
First, the \textbf{biased} situation, was a square landscape as above,
but with mean offspring location one tenth of a standard deviation up and to the right of the parent's location.
% (or 0.05\% of the edge length of the landscape)
This produces reverse-time gene flow down and to the left.
Second, we simulated a \textbf{valley} in a landscape with a $3 \times 5$ aspect ratio as above, 
in which fecundity of individuals in the middle 1/4 of the landscape was reduced by 50\%.
This produces gene flow ``uphill'', away from the ridge,
and thus an effective barrier across the ridge.
Third, we simulated a recent \textbf{expansion},
in which the middle 4/5 of the landscape was uninhabitable 400,000 generations,
until 500 generations before the present day.
The resulting expansion produces gene flow back away from the center,
similar to the ``valley''.

This results in reverse time gene flow down and to the left 
as in the biased asymmetric graph in Section \ref{sec:biased_migration}.
In terms of population genetics, 
this is very similar to the case where a population is expanding into new territory 
since in both cases, 
offspring tend to be further from the edge of the expansion than their parents.
The difference is that, in our case, 
the population is constrained to be a bounded distance 
away from the edge of the ``expansion''.
In both situations, individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift.
This results in substantially more noise in the isolation by distance plots 
as seen in Figure \ref{fig:ibd_comp}. 
The horizontal lines in that figure are likely from comparing genomes of individuals from
two families that have recently increased in size and geographical spread, 
resulting in many pairs of genomes with similar genetic distance 
but varying geographic distance.

As in Section \ref{sec:slim_unif}, we discretize the landscape into
both a $2 \times 2$ grid and a $4 \times 4$ grid for inference.
Looking at the medians of the inferred posteriors of the gene flow rates in the $2 \times 2$ case
(Figure \ref{fig:2x2grid_asb}),
it is immediately clear how strong the bias 
against reverse time gene flow out of the lower left corner is.
That is, hardly any individuals in the lower left area have ancestors
from other regions of the landscape.
This is to be expected as even though the single generation migration bias is small,
having a descendant in a location against the bias after a large number of generations
becomes less and less likely as the number of generations increases.
In the $4 \times 4$ discretization, 
the bottom left also shows very strongly biased gene flow (Figure \ref{fig:grid_bias_4x4_1}).
Most (but not all) of the other pairs also show bias in the expected direction.
Boxplots of the posterior distributions are shown in Figure \ref{fig:post_dists_bias_4x4_1}.
%\emph{maybe talk about/cite some biased random walk stuff here?}

%\emph{talk about some kind of log ratio test for the 4x4 case?}

\begin{figure}
\centering
 \includegraphics[scale=.35]{figs/ibd_comp}
\caption{Comparison of the isolation by distance plots for the square landscape with uniform migration
and the square landscape with biased migration.
Faster coalescence in the biased migration case results in more noticeable horizontal lines,
indicative of families that recently increased in size and geographical spread.}
\label{fig:ibd_comp}
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.5]{figs/2x2grid_asb}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $2 \times 2$ discretization of the square landscape with biased migration.
The reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in the bottom left of the landscape
have ancestors from other regions of the habitat,
consistent with migration biased up and to the right.}
\label{fig:2x2grid_asb}
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.8]{figs/grid_bias_4x4_1}
\caption{Medians of the posterior distributions of the inferred values of $g$ 
for the first replicate of the $4 \times 4$ discretization of the square landscape with biased migration.
As in the $2 \times 2$ case, 
the reverse time gene flow rates out of the bottom left section are very small,
indicating that very few individuals living in that area of the landscape
have ancestors from other regions of the graph.
Most, but not all of the other gene flow rates are biased 
in the direction we would expect.}
\label{fig:grid_bias_4x4_1}
\end{figure}

@@

With these methods, we provide a framework for
inferring demographic parameters of 
populations living on a habitat spread over a two dimensional landscape 
by using present day location and genetic sequencing information.
We are able to do inference of reverse time gene flow and coalescence rates 
with the matrix of mean coalescence times
in cases where the matrix is noise-free, noisy, or incomplete, 
or when it made after discretization of a simulated or real landscape.
To do this, we assume that the reverse time movements of lineages are Markov
and the population is near stationarity.
Typically, there is more error and uncertainty in the posterior distributions of the parameters
when there is more noise, slower coalescence rate, more missing values,
and multiple parameters for coalescence rate.

@@

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{figs/valley_locations}
    \includegraphics[width=0.45\textwidth]{figs/valley_eems}
    \includegraphics[width=0.45\textwidth]{figs/valley_commute_graph}
    \includegraphics[width=0.45\textwidth]{figs/valley_coal_graph}
    \includegraphics[width=0.45\textwidth]{figs/valley_commute_dists}
    \includegraphics[width=0.45\textwidth]{figs/valley_coal_dists}
    \caption{
        Results from the \textbf{valley} simulation:
        \textbf{(top left)} Positions of all individuals at the end of simulation.
        \textbf{(top right)} mean migration rates as estimated by EEMS.
        Below are shown
        posterior median migration rates 
            as estimated using 
            \textbf{(middle left)} the commute time method and
            \textbf{(middle right)} the the commute time method.
        Last are shown the posterior distributions, for
            \textbf{(bottom left)} the commute time method and
            \textbf{(bottom right)} the the commute time method.
        \label{sfig:valley_results}
    }
\end{figure}

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{figs/expansion_locations}
    \includegraphics[width=0.45\textwidth]{figs/expansion_eems}
    \includegraphics[width=0.45\textwidth]{figs/expansion_commute_graph}
    \includegraphics[width=0.45\textwidth]{figs/expansion_coal_graph}
    \includegraphics[width=0.45\textwidth]{figs/expansion_commute_dists}
    \includegraphics[width=0.45\textwidth]{figs/expansion_coal_dists}
    \caption{
        Results from the \textbf{expansion} simulation:
        \textbf{(top left)} Positions of all individuals at the end of simulation.
        \textbf{(top right)} mean migration rates as estimated by EEMS.
        Below are shown
        posterior median migration rates 
            as estimated using 
            \textbf{(middle left)} the commute time method and
            \textbf{(middle right)} the the commute time method.
        Last are shown the posterior distributions, for
            \textbf{(bottom left)} the commute time method and
            \textbf{(bottom right)} the the commute time method.
        \label{sfig:expansion_results}
    }
\end{figure}

@@

>>>>

\begin{abstract}
A number of methods commonly used in landscape genetics
use analogies to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical equivalence between electrical resistance between two nodes
and the ``commute time'',
which is the mean time for a random walk on that network to leave one node, visit the other, and return.
Here, we describe the differences between commute time
and coalescence time, which more accurately models genetic distances,
and explore the consequences for inference.
We implement a Bayesian method to infer effective movement rates and population sizes 
under both these models,
and find that inference using commute times
can dramatically mislead in the presence of biased gene flow.
Forwards-time simulation with continuous geography demonstrates that
coalescence-based inference remains more accurate than resistance-based methods on realistic data,
and highlights the need for methods that do not approximate continuous space
by a discrete set of populations.
\end{abstract}

@@

In this paper, we contrast 
coalescence time and resistance distance,
and develop an inference method for movement rates on a landscape
using both as estimated from genetic distances.
As we will see,
although the two methods are conceptually similar,
commute time inference may get the wrong answer if data derive from a coalescent process,
as does real genetic data, especially in the presence of asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{infer_resistance}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem --
inference of movement rates, in either a coalescence or commute time framework,
is an ill-conditioned problem,
implying the need for regularization to obtain reliable inference.
This explains the results of \citet{graves2013current},
who found a large area of resistance parameters
that produced equivalently good fits to genetic distance
(i.e., nonidentifiability, analogous to a flat likelihood surface).

Talk about MAPS.

@@

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.0 \citep{haller2017slim,haller2018treesequences},
with individuals living across continuous, two-dimensional geography (sometimes with barriers),
from which we recorded genomic data.
The basic simulation, which we modified to produce several other situations,
is as follows.
Each simulation had around $10^4$ diploid, hermaphroditic individuals, 
living across two-dimensional geography.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Simulations were done on either a square region with width equal to $80\sigma_d$,
or a $3:5$ rectangular region with the same area.
Offspring are dispersed away from the mother using the same kernel,
reflected from the boundaries of the habitable area.
In each simulation, $\sigma_d$ was chosen 
so that there were an average of 0.25 individuals per $\sigma_d^2$ units of area.
In order to prevent excessive population clumpiness \citep{felsensten1975pain}, 
we also model local competition. 
To do this, we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals;
then, the fitness of each individual is set to $\exp(-x/40)$, 
where $x$ is the sum of their interaction strengths with other individuals.
This reduces fitness for individuals in denser areas.
Individuals are chosen to reproduce proportional to their fitness.
Clumpiness is similar to what is seen in 
Figure \ref{fig:5x3b_continuous}.

Individuals have 10 pairs of ``chromosomes'' with $2 \times 10^5$ loci each 
and a recombination rate of $10^{-8}$ per generation per locus.
Neutral mutations were added at a rate of $10^{-9}$ per meiosis.
To make these simulations computationally feasible,
the mutation rate was set to 0.0 during the forwards simulation,
while tree sequences were recorded \citep{haller2018treesequences},
and mutations were added after the fact with msprime,
which is equivalent to including mutations during the forwards simulation 
but resulted in much faster run times \citep{kelleher2018xxx}.
In all cases, the number of mutations was more than adequate to reliably estimate genetic distance.

To compute genetic distances,
geographic space was partitioned into rectangular regions as depicted below.
within each region, 
individuals are sampled uniformly for ``genotyping'' 
from among those within the middle three quarters of each dimension of rectangle,
as shown for one situation in Figure \ref{fig:5x3b_continuous}.
This protocol was chosen as a compromise between sampling all individuals 
from near the center of each location,
which could result in a large number of close relatives being sampled, 
leading to an underestimate of the mean genetic distance for individuals in that location
and uniform sampling from the whole area of each grid, 
which may result in many sampled individuals being very close to other grid squares.
Mean genetic distances for each pair of regions is then computed as described above
from genome sequence provided by SLiM.
To estimate uncertainty, we computed the standard error of these mean distances
using the minumum number of individuals in each location.

@@

\subsection*{Code availability}

All code used for this work is available at \url{https://github.com/elundgre/gene-flow-inference}
(including an R package to perform inference)
and \url{https://github.com/petrelharp/isolation_by_coalescence} (including the SLiM simulations).

@@

The data we have used thus far are produced ``under the model'',
and should provide an accurate depiction of our method applied to discrete, randomly mating populations
whose connections by migration are known.
However, this is nearly always a rough approximation to reality,
in which organisms are distributed across continuous geography,
and ``populations'' are constructed by necessity, often driven by sampling locations.
We also applied coalescence and resistance inference,
as well as the popular resistance-based method EEMS \citep{petkova_eems},
to data from simulations on continuous geography.

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
each observed, empirical mean estimated using a sample of genotyeps
will deviate from this depending on both the individuals sampled (``sampling noise'')
and the stochasticity of the population history itself (``process noise'').
For more discussion of this latter contribution in other contexts, see
\citet{wakeley2012genealogies}, \citet{waples2009modeling}, or \citet{ralph2015empirical}.
In populations living across continuous geography,
both sources of noise can be substantially larger due to randomness in spatial locations.
This is hinted at by Figure \ref{fig:5x3b_continuous},
where we see that the distribution of individuals across the landscape is fairly patchy:
a barrier might be inferred in a region that has not happened to have a local population recently
only because of demographic stochasticity.
This might make inference using a discretization of continuous space considerably more difficult.

@@

\begin{figure}
\centering
    \includegraphics[width=0.8\textwidth]{figs/h_val_comp}
\caption{
    Comparison of the values of the genetic distance matrix $C$
    for a $4 \times 4$ discretization of a square landscape with uniform migration
    calculated from various situations.
    All values are plotted against the corresponding values from the $C$ matrix
    calculated using the first set of individuals from the first simulation,
    along with the $y=x$ line.
    Black circles are from a second (non-overlapping) set of individuals from the first simulation,
    red triangles are from the second simulation, 
    and blue plusses are for the third simulation.
    Mean absolute value relative differences between the new values of $C$ for each situation
    and the original values from the first set of individuals from the first simulation are
    1.8\% for the second set of indivuals from the first simulation,
    13.2\% for the second simulation, 
    and 7.3\% for the third simulation.
    } \label{fig:h_val_comp}
\end{figure}

Although we found substantial process noise
with a large, flat landscape with uniform, unbiased migration,
many biological processes are expected to increase between-replicate stochasticity.
To demonstrate this, we simulated a square, flat landscape as above,
but with \emph{biased migration}: 
each offspring's location was chosen randomly as before,
but with mean location $\sigma_d/10$
up and to the right of the parent's location.
% (or 0.05\% of the edge length of the landscape)
This produces reverse-time gene flow down and to the left.
As a population genetics model,
this case is similar to a population expansion,
replacing the leading edge of the expansion with the corner of landscape 
that migration bias leads away from.
In this situation, as in an expansion, 
individuals at the edge may have disproportionately more offspring,
leading to low genetic diversity and strong drift \citep{surfing}.
This results in substantially more noise in the data,
clearly visible in the isolation by distance plots of Figure \ref{fig:ibd_comp}. 
Despite this large contribution of process noise,
coalescence time inference correctly infers the bias on a broad scale:
using both a $2 \times 2$ and a $4 \times 4$ grid,
inferred gene flow rates down and to the left are consistently much larger
than those up and to the right 
(Supplemental Figures \ref{fig:2x2grid_asb}, \ref{fig:grid_bias_4x4_1}, 
and \ref{fig:post_dists_bias_4x4_1}).
However, inferred rates show substantial heterogeneity,
likely due to stochasticity of recent demography.

\begin{figure}
\centering
     \includegraphics[width=0.75\textwidth]{figs/ibd_comp}
    \caption{
        Genetic distance against geographic distance,
        for the square landscape with 
        \textbf{(above)} uniform migration
        and \textbf{(below)} biased migration.
        The horizontal lines in the latter plot
        are likely from comparing genomes of individuals from
        families that have recently increased in size and geographical spread, 
        resulting in many pairs of genomes with similar genetic distance 
        but varying geographic distance.
    \label{fig:ibd_comp}
    }
\end{figure}

\paragraph{Identifying a barrier}
We now revisit the earlier situation where a landscape has locations that are barriers to gene flow.
The landscape is a continuous version of this situation,
shown in Figure \ref{fig:5x3b_continuous},
where red bars depict uninhabitable regions that block migration.
Since the distribution of offspring locations is truncated at three standard deviations
away from the mother, a sufficiently thick uninhabitable area completely stops migration
directly across it.
We calculate mean genetic distances using 50 individuals 
from each grid location,
here giving us a total of 750 individuals.
The gene flow rates across the barriers are reliably estimated to be very small,
while other values have substantial variation:
in particular, gene flow rates are estimated to be low
across some edges that do not cross barriers;
this may be because of recent demographic stochasticity.
The barriers were also identified with comparable accuracy by both our resistance distance method
and EEMS, as shown in Supplemental Figure \ref{sfig:barrier_results}.
Note also that this analysis began by grouping samples into discrete ``populations'' appropriately;
in absence of a good \textit{a priori} idea of where barriers are,
this step might not be straightfoward.
% For instance, we may wonder whether a river or other geographical feature
% restricts gene flow.
% We would then discretize the landscape such that individuals on opposite sides of the 
% potential obstruction are in separate grid spaces.

\begin{figure}
\centering
     \includegraphics[scale=.8]{figs/ind_locs_5x3b_1}
     \includegraphics[height=0.3\textheight]{figs/grid_5x3b_1}
     \includegraphics[height=0.3\textheight]{figs/post_dists_5x3b_1_new}
    \caption{
        \textbf{(top)}
        Locations of all and selected individuals on the $5 \times 3$ landscape with barriers,
        with individuals used to compute the mean genetic distance matrix marked in bold.
        The red bars show the locations of uninhabitable regions of the landscape.
        They are sufficiently thick so that no migration can occur directly across them.
        \textbf{(bottom left)}
        Posterior medians of movement rates inferred using coalescence time inference,
        and \textbf{(bottom right)}
        boxplots of corresponding posterior distributions,
        with movement rates crossing the barriers colored in red.
        The grouping of individuals into a $5 \times 3$ grid is shown above,
        and labels are the same as in Figure \ref{fig:5x3b_grid}.
    \label{fig:5x3b_continuous}
    }
\end{figure}

\paragraph{Other barriers: population sinks and recent expansions}
We simulated two other situations on a continuous landscape that produced an bias in effective migration.
First, we simulated a \textbf{valley} in a landscape with a $5 \times 3$ aspect ratio as above, 
in which fecundity of individuals in the middle 1/4 of the landscape was reduced by 50\%
(a ``source-sink'' landscape).
Since individuals in the valley are more likely to have parents outside of the valley,
this produces gene flow ``uphill'', away from the ridge.
Next, we simulated a recent \textbf{expansion},
in which the middle 4/5 of the landscape was uninhabitable for 400,000 generations,
until 500 generations before the present day
(a model of ``secondary contact'').
The resulting expansion produces gene flow back away from the center,
similar to the ``valley''.

Both coalescence time inference and the resistance landscape inferred by EEMS
show a barrier to gene flow in the middle of the landscape (Figure \ref{fig:more_barriers}):
this is arguably correct, 
although there is in fact no current barrier to migration in either simulation.
Coalescence time inference also correctly infers 
the general bias in recent gene flow away from the center,
although as in other continuous-space simulations, 
there was substantial noise in the inferred patterns.
The resistance landscapes as inferred by EEMS show patchy, vertical bands of high resistance:
these patterns are commonly seen when attempting to use EEMS, 
whose underlying model is inherently symmetric, 
to describe asymmetric gene flow.
Our analysis above shows that this is an unavoidable consequence of the resistance approximation,
and our implementation of commute time inference,
which allows asymmetric gene flow, shows simliar patterns
(Supplemental Figures \ref{sfig:valley_results} and \ref{sfig:expansion_results}).

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{figs/valley_eems}
    \includegraphics[width=0.45\textwidth]{figs/valley_coal_graph}
    \includegraphics[width=0.45\textwidth]{figs/expansion_eems}
    \includegraphics[width=0.45\textwidth]{figs/expansion_coal_graph}
    \caption{
        \textbf{(Left:)}
        resistance landscapes inferred by EEMS
        and 
        \textbf{(right:)}
        gene flow rates inferred using coalescence time inference,
        from individual-based simulations on a continuous landscape,
        with \textbf{(top)} a \emph{valley} of reduced fecundity in the middle, and
        \textbf{(bottom)} a recent \emph{population expansion} from both edges of the range
        that met in the middle.
        The color scale shows the log migration rates inferred by EEMS.
        \label{fig:more_barriers}
    }
\end{figure}

@@

In this paper, we have studied the problem of
inferring demographic parameters of populations spread over heterogeneous two-dimensional landscapes
by using genetic and geographic distances between present-day samples.
In particular, we have shown that the resistance distance approximation
-- which underlies several of the most commonly-used tools of landscape genetics --
can produce misleading results, especially in the presence of biased gene flow.
We implement an alternative method, 
which uses coalescence times instead of the resistance distance approximation,
which provides good estimates of gene flow in a wider range of situations.
This weakness of resistance-based methods is not surprising --
the original papers describing Circuitscape \citep{mcrae_circuitscape} and EEMS \citep{petkova_eems}
present the resistance approximation as a necessary step for computational feasibility --
but perhaps not widely appreciated.

Our new method infers effective population sizes and gene flow rates quite well 
-- and, being Bayesian, provides estimates of uncertainty --
given data from discrete populations connected by migration.
Why do we not advocate our method as an immediate replacement for resistance-based methods?
The substantial uncertainty we saw on graphs with only tens of nodes
is indicative of a larger problem we face for realistic models.
There is a high degree of unavoidable uncertainty,
and process noise can lead to overfitting of models.
To produce reproducible, reliable results we need to address these problems.

We are able to do inference of reverse time gene flow and coalescence rates 
with the matrix of mean coalescence times
in cases where the matrix is noise-free, noisy, or incomplete, 
or when it made after discretization of a simulated or real landscape.
To do this, we assume that the reverse time movements of lineages are Markov
and the population is near stationarity.
Typically, there is more error and uncertainty in the posterior distributions of the parameters
when there is more noise, slower coalescence rate, more missing values,
and multiple parameters for coalescence rate.

@@

\begin{figure}
\centering
 \includegraphics[scale=.5]{figs/2x2grid_asb}
\caption{
    Medians of the posterior distributions of the inferred values of $g$ 
    for the first replicate of the $2 \times 2$ discretization of the square landscape with biased migration.
    The reverse time gene flow rates out of the bottom left section are very small,
    indicating that very few individuals living in the bottom left of the landscape
    have ancestors from other regions of the habitat,
    consistent with migration biased up and to the right.
    These inferred rates make it clear how strong the bias 
    against reverse time gene flow out of the lower left corner is.
    Hardly any individuals in the lower left area have ancestors
    from other regions of the landscape.
    This is to be expected as even though the single generation migration bias is small,
    having a descendant in a location against the bias after a large number of generations
    becomes less and less likely as the number of generations increases.
    \label{fig:2x2grid_asb}
    }
\end{figure}

\begin{figure}
\centering
 \includegraphics[scale=.8]{figs/grid_bias_4x4_1}
\caption{
    Medians of the posterior distributions of the inferred values of $g$ 
    for the first replicate of the $4 \times 4$ discretization 
    of the square landscape with biased migration.
    As in the $2 \times 2$ case, 
    the reverse time gene flow rates out of the bottom left section are very small,
    indicating that very few individuals living in that area of the landscape
    have ancestors from other regions of the graph.
    Most, but not all of the other gene flow rates are biased 
    in the direction we would expect.
    Boxplots of the posterior distributions are shown in 
    Supplemental Figure \ref{fig:post_dists_bias_4x4_1}.
    }
\label{fig:grid_bias_4x4_1}
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{figs/valley_locations}
    \includegraphics[width=0.45\textwidth]{figs/valley_eems}
    \includegraphics[width=0.45\textwidth]{figs/valley_commute_graph}
    \includegraphics[width=0.45\textwidth]{figs/valley_coal_graph}
    \includegraphics[width=0.45\textwidth]{figs/valley_commute_dists}
    \includegraphics[width=0.45\textwidth]{figs/valley_coal_dists}
    \caption{
        Results from the \textbf{valley} simulation:
        \textbf{(top left)} Positions of all individuals at the end of simulation.
        \textbf{(top right)} mean migration rates as estimated by EEMS.
        Below are shown
        posterior median migration rates 
            as estimated using 
            \textbf{(middle left)} the commute time method and
            \textbf{(middle right)} the coalescent time method.
        Last are shown the posterior distributions, for
            \textbf{(bottom left)} the commute time method and
            \textbf{(bottom right)} the coalescent time method.
        \label{sfig:valley_results}
    }
\end{figure}

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{figs/expansion_locations}
    \includegraphics[width=0.45\textwidth]{figs/expansion_eems}
    \includegraphics[width=0.45\textwidth]{figs/expansion_commute_graph}
    \includegraphics[width=0.45\textwidth]{figs/expansion_coal_graph}
    \includegraphics[width=0.45\textwidth]{figs/expansion_commute_dists}
    \includegraphics[width=0.45\textwidth]{figs/expansion_coal_dists}
    \caption{
        Results from the \textbf{expansion} simulation:
        \textbf{(top left)} Positions of all individuals at the end of simulation.
        \textbf{(top right)} mean migration rates as estimated by EEMS.
        Below are shown
        posterior median migration rates 
            as estimated using 
            \textbf{(middle left)} the commute time method and
            \textbf{(middle right)} the coalescence time method.
        Last are shown the posterior distributions, for
            \textbf{(bottom left)} the commute time method and
            \textbf{(bottom right)} the coalescence time method.
        \label{sfig:expansion_results}
    }
\end{figure}
